---
title: "Geospatial Poisson in R"
output: html_notebook
---

```{r}
library(tidyverse)
library(sf)
```

### Part 1 - Data Preparation

Perform a spatial join to determine the census tract in which each burglary occurred. Hint: You may want to make use of the st_as_sf function in order to convert the burglaries data into an sf object.

After performing the spatial join, merge in the census data. Note: Make sure that the final dataset contains all census tracts.


```{r}
burglaries <- read_csv("../geospatial-poisson-ndidiox/data/burglaries_2023.csv")
```

```{r}
dc_shape <- st_read("../geospatial-poisson-ndidiox/data/DC/DC.shx")
```

```{r}
dc_shape |>
  ggplot() +
  geom_sf()
```


```{r}
burglaries_geo <- st_as_sf(burglaries |> drop_na('latitude'),
                           coords = c('longitude', 'latitude'),
                           crs = st_crs(dc_shape)
                           )
```


```{r}
dc_shape |>
  ggplot() +
  geom_sf() +
  geom_sf(data = burglaries_geo, size = 0.1)
```

```{r}
burglaries_dc <- st_join(burglaries_geo, dc_shape, join = st_within)
```


```{r}
census <- read_csv("C:/Users/ndidi/Documents/NSS_Projects/geospatial-poisson-ndidiox/data/census.csv")
```

```{r}
final_df <- left_join(burglaries_dc, census, by = join_by(TRACTCE == tract))
```


```{r}
final_df <- select(final_df, -c(primary_key, STATEFP, COUNTYFP, NAME.x, NAME.y, state, county, NAMELSAD))
```

```{r}
final_df <- final_df |>
  filter(median_income >= 0)
```



```{r}
final_df |> head()
```

### Part 2 - Exploratory Analysis
Perform some exploratory analysis on your prepared dataset.

Aggregate the data by census tract. Warning: each incident can appear multiple times if there are multiple victims, so be sure that you aren't double-counting any incidents.

```{r}
final_df <- final_df |>
  group_by(TRACTCE) |>
  mutate(num_burglaries = n_distinct(incident_number))
```


# Which census tract had the highest number of burglaries?

```{r}
burg_per_tract <- final_df |>
  st_drop_geometry() |>
  group_by(TRACTCE) |>
  summarise(burglaries = n_distinct(incident_number)) |>
  arrange(desc(burglaries))

burg_per_tract
```
# Tract 016000 had the highest number of burglaries


# Which census tract had the highest number of burglaries per 1000 residents?

```{r}
per_1000 <- final_df |>
  st_drop_geometry() |>
  group_by(TRACTCE) |>
  summarise(burg_per_1000 = (n_distinct(incident_number) / unique(population)) * 1000) |>
  arrange(desc(burg_per_1000)) |>
  ungroup()

per_1000
```
# Tract 016000 had the highest number of burglaries per 1000 residents.

```{r}
# add num_burg_per_1000 column to dataframe
final_df <- final_df |>
  group_by(TRACTCE) |>
  mutate(num_burg_per_1000 = (n_distinct(incident_number) / unique(population)) * 1000)
```


We're interested in the relationship between median income and number of aggravated burglaries, so examine those variables on their own and together to see what you can find. You may want to perform additional calculations, create plots, etc.

```{r}
final_df |>
  ggplot(aes(x = median_income, y = num_burglaries )) +
  geom_point()
```

```{r}
final_df |>
  ggplot(aes(x = median_income, y = num_burg_per_1000 )) +
  geom_point()
```

```{r}
final_df |>
  ggplot(aes(x = population)) +
  geom_histogram()
```


### Part 3 - Statistical Modeling

Fit a Poisson regression model with target variable the rate of burglaries per census tract and with predictor the median income. Offset using the log of the population so that we are looking at the rate of burglaries per population instead of the number of burglaries. How can you interpret the meaning of the output? How do the estimates from the model compare to the observed data?

```{r}
pr_burg <- glm(num_burglaries ~ median_income,
               data = final_df,
               family = poisson)
```


```{r}
pr_burg_offset <- glm(num_burglaries ~ median_income + offset(log(population)),
               data = final_df,
               family = poisson)
```


```{r}
```


```{r}
```


